Gravitational field calculation in spherical coordinates using variable densities in depth

### 2018-05-08 manuscript/manuscript.tex
Added empirical grid search figure in manuscript and more editing

@@

It worth mention that either a greater $D$ or a lower $\delta$ 
increases the number of discretizations of the tesseroid.
Although a greater $D$ may divide it in the latitude, longitude 
and/or radial direction; a lower $\delta$ will only discretize the 
original tesseroid in the radial direction.
This means that a decrease in $\delta$ will produce a more efficient 
computation than an increase in $D$.
\todo[inline]{Should we keep this paragraph? The same is explained bellow.}

@@

On the other hand, we have no a priori information about the numerical 
error of points with $\delta > \delta_p$ (region C on 
Figure \ref{fig:D-delta-grid-search}).
Although, if we find a point with acceptable numerical error in this 
region, we know it will produce more discretizations than the 
($D_\text{linear}$, $\delta_p$) point, leading to a non-efficient 
computation.
This is due to the fact that even an increase in $\delta$ will produce 
less density-based discretizations, the increase in $D$ will introduce 
even more subdivisions not only in the radial direction, but 
also in the latitude and longitude ones.

@@

In order to determine the value of $\delta_p$ we should compute the 
differences between the numerical model and the analytical solution for 
a spherical shell with an exponential density.
Thus we must obtain this analytical solution and define a tesseroid 
model to numerically approximate the gravitational fields.

Firstly, the analytical solution of the gravitational fields generated 
by a spherical shell with a density functions as the one in equation 
\ref{eq:density-exp} can be computed as the sum of the fields generated 
by a spherical shell with constant density $\rho(r') = C$ and another 
shell with an exponential density like:

\noindent where $A$, $\Delta h$ and $b$ are constants.
The gravitational potential field generated by a spherical shell with 
this density on any external point at a distance $r$ from its centre 
can be calculated by solving the integral in equation \ref{eq:shell-pot},
resulting as follows:

Secondly, we proposed a thin and a thick spherical shells with 1km and 
35km of thickness, respectively, modelled by a mesh of $30^\circ \times 
30^\circ$ tesseroids.
We have assigned the exponential density function proposed on equation 
\ref{eq:density-exp} to each spherical shell model, choosing some values 
of the constant $b$ in order to test the numerical approximation for 
different density variations.

Finally, we computed the gravitational potential ($V$), the 
vertical component of its gradient ($g_z$) and the $g_zz$ component of 
the Marussi tensor on the big grid (see Table \ref{tab:grids}) by 
keeping the distance-size ratio equal to the $D$ values obtained in the 
density case for its corresponding gravitational field and exploring 
different values of $\delta$.
We then computed the percentage difference between this numerical 
results and the analytical solutions.

>>>>

The analytical solution of the gravitational fields generated 
by a spherical shell with a density functions as the one in equation 
\ref{eq:density-exp} can be computed as the sum of the fields generated 
by a spherical shell with constant density $\rho(r') = C$ and another 
shell with an exponential density like:

\noindent where $A$, $\Delta h$ and $b$ are constants.
The gravitational potential field generated by a spherical shell with 
the latter density on any external point at a distance $r$ from its centre 
can be calculated by solving the integral in equation \ref{eq:shell-pot},
resulting as follows:

Then, we proposed a thick spherical shell with a thickness of 35km and 
modelled by a mesh of $30^\circ \times 30^\circ$ tesseroids.
We have assigned the density function proposed on equation 
\ref{eq:density-exp} to our spherical shell model with a $b$ value of 
1km, i.e. a highly varying density function.

We computed the gravitational potential ($V$), the vertical component 
of its gradient ($g_z$) and the $g_zz$ component of the Marussi tensor 
on the big grid (see Table \ref{tab:grids}) by exploring different 
values of $D$ and $\delta$.
These results were used to obtain the percentage difference between 
them and their corresponding analytical solutions.

On Figure \todo[inline]{referencia a figura grid search} we show 
the resulting percentage difference for each gravitational field.
The points inside the dotted lines are the ones that present a 
numerical error lower than the 0.1\% threshold, i.e. they represent the 
eligible ($D$, $\delta$) pairs.
The one that produces the more efficient computation must be chosen as 
the default $D$ and $\delta$ value.

\begin{figure}
\centering
\includegraphics[width=\linewidth]
    {figures/grid-search.pdf}
\caption{
    }
\label{fig:grid-search}
\end{figure}

@@

%~ It worth mention that either a greater $D$ or a lower $\delta$ 
%~ increases the number of discretizations of the tesseroid.
%~ Although a greater $D$ may divide it in the latitude, longitude 
%~ and/or radial direction; a lower $\delta$ will only discretize the 
%~ original tesseroid in the radial direction.
%~ This means that a decrease in $\delta$ will produce a more efficient 
%~ computation than an increase in $D$.
%~ \todo[inline]{Should we keep this paragraph? The same is explained bellow.}

@@

On the other hand, we have no a priori information about the numerical 
error of points with $\delta > \delta_p$ (region C on 
Figure \ref{fig:D-delta-grid-search}).
Nevertheless, the numerical error exploration shown on
Figure \todo[inline]{ref a figura grid search} doesn't present any 
point in this region with error lower than the threshold.
Even if we are analysing a single case, it proves that no point in 
region C is eligible as default ($D$, $\delta$) pair because it should 
ensure numerical accuracy for every exponential density scenario.

We can arrive to the same conclusion by analysing the computation 
efficiency: even if we find a point with acceptable numerical error in 
this region, we know it will produce more discretizations than the 
($D_\text{linear}$, $\delta_p$) point, leading to a non-efficient 
computation.
This is due to the fact that even an increase in $\delta$ will produce 
less density-based discretizations, the increase in $D$ will introduce 
even more subdivisions not only in the radial direction, but 
also in the latitude and longitude ones.

@@

%~ In order to determine the value of $\delta_p$ we should compute the 
%~ differences between the numerical model and the analytical solution for 
%~ a spherical shell with an exponential density.
%~ Thus we must obtain this analytical solution and define a tesseroid 
%~ model to numerically approximate the gravitational fields.

%~ Secondly, we proposed a thin and a thick spherical shells with 1km and 
%~ 35km of thickness, respectively, modelled by a mesh of $30^\circ \times 
%~ 30^\circ$ tesseroids.
%~ We have assigned the exponential density function proposed on equation 
%~ \ref{eq:density-exp} to each spherical shell model, choosing some values 
%~ of the constant $b$ in order to test the numerical approximation for 
%~ different density variations.

%~ Finally, we computed the gravitational potential ($V$), the 
%~ vertical component of its gradient ($g_z$) and the $g_zz$ component of 
%~ the Marussi tensor on the big grid (see Table \ref{tab:grids}) by 
%~ keeping the distance-size ratio equal to the $D$ values obtained in the 
%~ density case for its corresponding gravitational field and exploring 
%~ different values of $\delta$.
%~ We then computed the percentage difference between this numerical 
%~ results and the analytical solutions.
